---
title: "RBerkeley functions demonstration"
author: "Abhinav Agarwal"
date: "11 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This document describes the functions supported by RBerkeley API for BerkeleyDB. This is also my submission for GSoC 2019 project, aimed at bringing RBerkeley back to CRAN Repository. 
All the functions are listed in alphabetical order, with their demo usage.
No description has been provided for any function, as it would be more appropriate to refer to BerkeleyDB C API for that.

Loading the API:
```{r}
library(RBerkeley)
db_version()
```


### BerkeleyDB Database Operations and Methods

#### db_create
```{r}
dbh<-db_create()
dbh
```

#### db_open
Mask for DB->open() in C API
```{r}
db_open(dbh)
```

#### db_put
Mask for DB->put()	in C API
```{r}
db_put(dbh, key="abc", data="def")
db_put(dbh, key="123", data="456") #Automatically serialized by R

```

#### db_get
Mask for DB->get() in C API
```{r}
db_get(dbh, key="abc") #Output in serialized format
unserialize(db_get(dbh, key="abc")) #Output in unserialized format
```

#### db_compact
Mask for DB->compact() in C API. Unfortunately, there seems to be some error in the implementation of this function, as R session crashes everytime I try to execute it. The function is not executed here.
```{r eval=FALSE}
db_compact(dbh)
```

#### db_del & db_exists

Mask for Db::del() & DB->exists() in C API
 
```{r}
db_strerror(db_exists(dbh, key="abc"))
db_del(dbh, key="abc")
db_strerror(db_exists(dbh, key="abc"))
```

Note: While db_fd() has been mentioned in Documentation, it has not yet been implemented in RBerkeley.

#### db_get_byteswapped
Maks for Db::get_byteswapped in C API
```{r eval=FALSE, include=FALSE}
db_get_byteswapped((dbh)) #Error in db_get_byteswapped((dbh)) : invalid 'db' handle
```
Possibly, implementation of this function is missing `as.DB(dbh)`.

#### db_get_type	

Mask for DB->get_type in C API

```{r}
db_get_type(dbh)
```

#### db_key_range
Mask for  DB->key_range in C API.

```{r}
db_key_range(dbh,key=123)
```
#### db_rename
Mask for DB->rename in C API.

```{r}
db_get_dbname(dbh)

db_rename(dbh,"access.db","","gsoc")
db_get_dbname(dbh)

```


#### db_cursor
Mask for DB->cursor() in C API.

```{r}
dbc <- db_cursor(dbh)
dbc
```
#### dbcursor_put
Mask for DBcursor->put() in C API.
```{r}
dbcursor_put(dbc, key=1000L, data=50L, flags=mkFlags(DB_KEYFIRST))
```


#### dbcursor_count
Mask for DBcursor->count() in C API.
```{r}
dbcursor_count(dbc)
```

#### dbcursor_get
Mask for DBcursor->get() in C API
```{r}
rs<-dbcursor_get(dbc, key=100,flags=mkFlags(DB_FIRST))
lapply(rs[[1]], unserialize)
```

Note: The documentation erroneously mentions `dbcursor_pget` as a method. No such function exists in the API.

#### dbcursor_del
Mask for DBcursor->del() in C API


```{r}
dbcursor_del(dbc)
dbcursor_get(dbc, key=100,flags=mkFlags(DB_SET)) #null
```
#### dbcursor_dup
Mask for DBcursor->dup() in C API
```{r}
dbc2<-dbcursor_dup(dbc)
dbc2
```

#### dbcursor_set_priority & dbcursor_get_priority
Mask for DBcursor->set_priority() and DBcursor->get_priority() respectively in C API

```{r}
dbcursor_set_priority(dbc,as.integer(5))
dbcursor_get_priority(dbc)

```
Note: dbcursor_get_priority is not mentioned in API documentation

#### dbcursor_close
Mask for DBcursor->close() in C API

```{r}
dbcursor_close(dbc)
```


#### db_stat

Mask for DB->stat in C API

```{r}
db_stat(dbh)
```
`db_sync`, `db_upgrade` and `db_truncate` are some other functions which can be implemented in a similar fashion.

#### db_close
Mask for DB->close() in C API
```{r}
db_close(dbh)
dbh #Error 
```
Similarly, `db_rename` and `db_remove` can be used for relevant tasks. 


### Conclusion

RBerkeley is a very powerful library, and I would work as hard as I can to see it back on CRAN by addressing the relevant issues.

### References

* RBerkeley documentation and Vignette
* Oracle Berkeley DB C API
